### 2014-08-18 analysis/context_switching.Rmd
pre-proof-read version of the paper added; some correlation analysis added; other old analysis corrected

Context Switching
=================

In this set of experiments we analyse a new multi-context model with a binary opinion dynamics game. In this case, the difference between the context switching model and the context permeability one is that agents no longer interact in the multiple  networks at the same time (being able to select any neighbour at a given simulation step) but rather become active in one network at a time. Agents are also able to switch to a different network at the end of each interaction with a partner. The agent switches based on a probability associated with the current network in which it is currently active.

@@

We then freeze the switching probabilities and vary the number of networks. We compare the results with the context permeability in terms of percentage of convergences to consensus and average number of encounters.

@@

The analysis for each context switching probability span is as follows: 
```{r load-parameters}
analyse_contour <- function(paramfile, encountersfile, net1label, net2label){
  #load some utilities
source("../R/experiment_utils.R")

#Read the Parameters for this experiment
params <- read.exp.parameters(param_file_name=paramfile)

#read data for number of encounters

require(data.table)

encounter_data <- fread(encountersfile)
encounter_data <- as.data.frame(encounter_data)[,-ncol(encounter_data)]
head(encounter_data)

#merge with configuration context switching values
require(reshape2)
encounter_data <- merge(encounter_data, params[,c("cfg.id","network.0.cs","network.1.cs")], by="cfg.id")
head(encounter_data)

#melt data to use cast
melted_encounters <- melt(encounter_data,id.vars=c("cfg.id","run","network.0.cs","network.1.cs"), measure.vars=c("total-encounters"))
head(melted_encounters)

@@

dev.off() 
print(encouters_contour)
gc()
}

analyse_persp <- function(paramfile, encountersfile,output_prefix,net1label,net2label){
    #load some utilities
source("../R/experiment_utils.R")

#Read the Parameters for this experiment
params <- read.exp.parameters(param_file_name=paramfile)

#read data for number of encounters

require(data.table)

encounter_data <- fread(encountersfile)
encounter_data <- as.data.frame(encounter_data)[,-ncol(encounter_data)]
head(encounter_data)

#merge with configuration context switching values
require(reshape2)
encounter_data <- merge(encounter_data, params[,c("cfg.id","network.0.cs","network.1.cs")], by="cfg.id")
head(encounter_data)

#melt data to use cast
melted_encounters <- melt(encounter_data,id.vars=c("cfg.id","run","network.0.cs","network.1.cs"), measure.vars=c("total-encounters"))
head(melted_encounters)
  #reshape the data to display average number of encounters
encounter_table <- dcast(data=melted_encounters, network.0.cs~network.1.cs, mean)

@@

zlab<-"\n\n\n Log(Avg. Encounters)"
pdf(paste("./pdf/",output_prefix,"_switching_log_persp.pdf"), width=5, height=5)
plot_persp_span(x = seq(0,1,0.05), y = seq(0,1,0.05),data_matrix = log(encounter_matrix), breaks=log(c(1:10)),xlab=xlab,ylab=ylab,zlab=zlab,pallete="PRGn")
dev.off()
plot_persp_span(x = seq(0,1,0.05), y = seq(0,1,0.05),data_matrix = log(encounter_matrix), breaks=log(c(1:10)),xlab=xlab,ylab=ylab,zlab=zlab,pallete="PRGn")
}
```

@@

**Note** I don't need custom breaks for the **persp plots**, its simpler to apply a logarithm to a set of n numbers, normalize them inside the plotting function, and multiply this by the maximum value in the dataset, this gets me the effect I'm looking for both the filling and contour lines.

@@

@@

@@

## Comparison for K-Regular Networks ##
```{r}
#**************************************************************************
#   read encounter results for context permeability (for comparison)
#**************************************************************************
params <- "../data/context_permeability/regular_networks/param-space_2014-05-30_16:47:53experiment:0_.csv"
encounters <- "../data/context_permeability/regular_networks/num_encounters.csv"

@@

require(data.table)

encounter_data <- fread(encounters)
encounter_data <- as.data.frame(encounter_data)[,-ncol(encounter_data)]
#merge with configuration context switching values
require(reshape2)
encounter_data <- merge(encounter_data, params[,c("cfg.id","network.0.k","num.networks")], by="cfg.id")
head(encounter_data)
#melt data to use cast
melted_encounters <- melt(encounter_data,id.vars=c("cfg.id","run","network.0.k","num.networks"), measure.vars=c("total-encounters"))
head(melted_encounters)
permeability_encounters <- melted_encounters

#***************************************************************
#   read encounter results for context switching
#***************************************************************
params <- "../data/context_switching/multiple_networks/regular_networks/param-space_2014-06-22_00:33:19experiment:10_.csv"

@@

require(data.table)

@@

#merge with configuration context switching values
require(reshape2)
encounter_data <- merge(encounter_data, params[,c("cfg.id","network.0.cs","network.0.k","num.networks")], by="cfg.id")
head(encounter_data)

@@

@@

  #gets too messy with permeability data 
  #perm_data <- dcast(perm_encounters, network.0.k + num.networks ~ variable, mean)
  #colnames(perm_data) <- c("k","num-networks", "avg-encounters")

@@

  #add line for permeability
  #plot <- plot +  geom_point(data=perm_data, aes(x=as.factor(perm_data$k), y=perm_data$"avg-encounters", shape="permeability", color=as.factor(perm_data$"num-networks")),size = 3.5)
  #plot +  geom_line(data=perm_data, aes(x=as.factor(perm_data$k), y=perm_data$"avg-encounters",group=as.factor(perm_data$"num-networks"),linetype=as.factor(plot_data$"num-networks"),color=as.factor(perm_data$"num-networks")),size=1)

@@

   print(plot)
}
gc()

@@

>>>>

Context Switching
=================
In this set of experiments we analyse a new multi-context model with a binary opinion dynamics game. In this case, the difference between the context switching model and the context permeability one is that agents no longer interact in the multiple  networks at the same time (being able to select any neighbour at a given simulation step) but rather become active in one network at a time. Agents are also able to switch to a different network at the end of each interaction with a partner. The agent switches based on a probability associated with the current network in which it is currently active.

@@

We then freeze the switching probabilities and vary the number of networks. We compare the results with the context permeability in terms of average number of encounters to achieve consensus.

We will start by loading the dependencies for this report.
```{r}
source("../R/experiment_utils.R")

library(data.table)
library(reshape2)
library(ggplot2)

@@

We will use the following procedures to construct the contour plots and perspective "3D" plots.
```{r load-parameters}
analyse_contour <- function(paramfile, encountersfile, net1label, net2label){
  #Read the Parameters for this experiment
  params <- read.exp.parameters(param_file_name=paramfile)

  #read data for number of encounters
  encounter_data <- fread(encountersfile)
  encounter_data <- as.data.frame(encounter_data)[,-ncol(encounter_data)]

  #merge with configuration context switching values
  encounter_data <- merge(encounter_data, params[,c("cfg.id","network.0.cs","network.1.cs")], by="cfg.id")

  #melt data to use cast
  melted_encounters <- melt(encounter_data,id.vars=c("cfg.id","run","network.0.cs","network.1.cs"), measure.vars=c("total-encounters"))

@@

  dev.off() 
  print(encouters_contour)
  gc()
}
```

```{r}
analyse_persp <- function(paramfile, encountersfile,output_prefix,net1label,net2label){
  #Read the Parameters for this experiment
  params <- read.exp.parameters(param_file_name=paramfile)

  #read data for number of encounters
  encounter_data <- fread(encountersfile)
  encounter_data <- as.data.frame(encounter_data)[,-ncol(encounter_data)]

  #merge with configuration context switching values 
  encounter_data <- merge(encounter_data, params[,c("cfg.id","network.0.cs","network.1.cs")], by="cfg.id")

  #melt data to use cast
  melted_encounters <- melt(encounter_data,id.vars=c("cfg.id","run","network.0.cs","network.1.cs"), measure.vars=c("total-encounters"))

  #reshape the data to display average number of encounters
  encounter_table <- dcast(data=melted_encounters, network.0.cs~network.1.cs, mean)

@@

  zlab<-"\n\n\n Log(Avg. Encounters)"
  pdf(paste("./pdf/",output_prefix,"_switching_log_persp.pdf"), width=5, height=5)
  plot_persp_span(x = seq(0,1,0.05), y = seq(0,1,0.05),data_matrix = log(encounter_matrix), breaks=log(c(1:10)),xlab=xlab,ylab=ylab,zlab=zlab,pallete="PRGn")
  dev.off()
  plot_persp_span(x = seq(0,1,0.05), y = seq(0,1,0.05),data_matrix = log(encounter_matrix), breaks=log(c(1:10)),xlab=xlab,ylab=ylab,zlab=zlab,pallete="PRGn")
}

@@

**Note** I don't need custom breaks for some **persp plots**, its simpler to apply a logarithm to a set of n numbers, normalize them inside the plotting function, and multiply this by the maximum value in the dataset, this gets me the effect I'm looking for both the filling and contour lines.

@@

### Scale-Free Network with D=5 and Scale-free Network with D=1 ###

```{r}
params <- "../data/context_switching/switching_span/scale-free d=1 scale-free d=5/param-space_2014-07-22_20:33:49experiment:12_.csv"
encounters <- "../data/context_switching/switching_span/scale-free d=1 scale-free d=5/num_encounters.csv"
output_prefix <- "context_switching_sf_d5_d1"

net1label <- "\n\nSwitching for Network 1 \n scale-free network, d=5"
net2label <- "\n\nSwitching for Network 2 \n scale-free network, d=1"

@@

analyse_contour(params,encounters,net1label,net2label)
```

@@

## Comparison for K-Regular Networks ##
```{r}
#**************************************************************************
#   read encounter results for context permeability (for comparison)
#**************************************************************************
params <- "../data/context_permeability/regular_networks/param-space_2014-07-23_11:40:44experiment:0_.csv"
encounters <- "../data/context_permeability/regular_networks/num_encounters.csv"

@@

encounter_data <- fread(encounters)
encounter_data <- as.data.frame(encounter_data)[,-ncol(encounter_data)]
#merge with configuration context switching values
encounter_data <- merge(encounter_data, params[,c("cfg.id","network.0.k","num.networks")], by="cfg.id")
head(encounter_data)
#melt data to use cast
melted_encounters <- melt(encounter_data,id.vars=c("cfg.id","run","network.0.k","num.networks"), measure.vars=c("total-encounters"))
head(melted_encounters)
permeability_encounters <- melted_encounters
```

Now we get the data for the k-regular networks with homogenous values of switching probability (the same probability in all the networks). --We have seen previously that for less connected networks 
simetry helps with the convergence: as such we take 0.25 0.5 and 0.75 as switching probabilities representing low, medium, and high switching respectively. 
```{r}
#***************************************************************
#   read encounter results for context switching
#***************************************************************
params <- "../data/context_switching/multiple_networks/regular_networks/param-space_2014-06-22_00:33:19experiment:10_.csv"

@@

@@

#merge with configuration context switching values
encounter_data <- merge(encounter_data, params[,c("cfg.id","network.0.cs","network.0.k","num.networks")], by="cfg.id")
head(encounter_data)

@@

  #export to pdf
  pdf(paste("./pdf/"," context_switching_encounters_comp_kregular_groupedbynets_",n,".pdf",sep = ""), width=7, height=5)
  print(plot)
  dev.off()

@@

@@

@@

  pdf(paste("./pdf/"," context_switching_encounters_comp_kregular_groupedbycs_",cs,".pdf",sep=""), width=7, height=5)
  print(plot)
  dev.off()
}
gc()
```

## Comparison for Scale-Free Networks ##

First, we get the permeability results for scale-free networks for comparison.
```{r}
#**************************************************************************
#   read encounter results for context permeability (for comparison)
#**************************************************************************
params <- "../data/context_permeability/scale-free_networks/param-space_2014-05-30_18:38:14experiment:1_.csv"
encounters <- "../data/context_permeability/scale-free_networks/num_encounters.csv"

encounter_data <- fread(encounters)
encounter_data <- as.data.frame(encounter_data)[,-ncol(encounter_data)]
#merge with configuration context switching values
encounter_data <- merge(encounter_data, params[,c("cfg.id","network.0.d","num.networks")], by="cfg.id")
head(encounter_data)
#melt data to use cast
melted_encounters <- melt(encounter_data,id.vars=c("cfg.id","run","network.0.d","num.networks"), measure.vars=c("total-encounters"))
head(melted_encounters)
permeability_encounters <- melted_encounters
```

We now create the plots to compare scale-free networks with context permeability. 
```{r}
#***************************************************************
#   read encounter results for context switching
#***************************************************************
params <- "../data/context_switching/multiple_networks/scale-free_networks/param-space_2014-06-22_00:30:57experiment:10_.csv"

encounters <- "../data/context_switching/multiple_networks/scale-free_networks/num_encounters.csv"

#merge with configuration context switching values
encounter_data <- merge(encounter_data, params[,c("cfg.id","network.0.cs","network.0.d","num.networks")], by="cfg.id")
head(encounter_data)

#melt data to use cast
melted_encounters <- melt(encounter_data,id.vars=c("cfg.id","run","network.0.cs","network.0.d","num.networks"), measure.vars=c("total-encounters"))
head(melted_encounters)

@@

```{r}
for(n in 2:5){
  #plot for 2 networks
  current_encounters<-melted_encounters[melted_encounters$"num.networks" == n,]
  perm_encounters<-permeability_encounters[permeability_encounters$"num.networks" == n,]

  plot_data <- dcast(current_encounters, network.0.d + network.0.cs ~ variable, mean)
  colnames(plot_data) <- c("d","cs","avg-encounters")

  perm_data <- dcast(perm_encounters, network.0.d ~ variable, mean)
  colnames(perm_data) <- c("d","avg-encounters")

  plot <- ggplot(plot_data, aes(x=as.factor(plot_data$d), y=plot_data$"avg-encounters")) 
  plot <- plot +  geom_point(aes(shape=as.factor(plot_data$cs), color=as.factor(plot_data$cs)),size = 3.5)
  plot <- plot +  geom_line(aes(color=as.factor(plot_data$cs), group=as.factor(plot_data$cs), linetype=as.factor(plot_data$cs)),size=1)   
  #add line for permeability baseline
  plot <- plot + geom_point(data=perm_data, aes(x=as.factor(perm_data$d), y=perm_data$"avg-encounters", shape="permeability", color="permeability"),size = 3.5)
  plot <- plot + geom_line(data=perm_data, aes(x=as.factor(perm_data$d), y=perm_data$"avg-encounters", group="Permeability", linetype="permeability", color="permeability"),size=1)

  plot <- plot + labs(color="CS", shape="CS", linetype="CS", y="Average Encounters", x="D",
                       title=paste("Scale-Free Networks With Number of Networks =",n))

  pdf(paste("./pdf/"," context_switching_encounters_comp_sf_groupedbynets_",n,".pdf", sep=""), width=7, height=5)
  print(plot)
  dev.off()

```{r}

  plot_data <- dcast(current_encounters, network.0.d + num.networks ~ variable, mean)
  colnames(plot_data) <- c("d","num-networks","avg-encounters")

  plot <- ggplot(plot_data, aes(x=as.factor(plot_data$d), y=plot_data$"avg-encounters")) 
  plot <- plot +  geom_point(aes(shape=as.factor(plot_data$"num-networks"), color=as.factor(plot_data$"num-networks")),size = 3.5)
  plot <- plot +  geom_line(aes(color=as.factor(plot_data$"num-networks"), group=as.factor(plot_data$"num-networks"), linetype=as.factor(plot_data$"num-networks")),size=1)

  plot <- plot + labs(color="Num. Networks", shape="Num. Networks", linetype="Num. Networks", y="Average Encounters", x="D",
                      title=paste("Scale-Free Networks With CS =",cs))

  pdf(paste("./pdf/"," context_switching_encounters_comp_sf_groupedbycs_",cs,".pdf",sep=""), width=7, height=5)
  print(plot)
  dev.off()
}
gc()
